Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-01-05 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

\reply:

@@

@@

\vspace{5mm}
\noindent \aecom \textbf{1 - Reply to AE comment 4}: 
T hank you for explaining the sampling limitations to me in your reply.  However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.

@@

@@

@@

>>>>

\reply: We agree and have addressed this point in \aecom \textbf{1 - Reply to AE comment 4}.

@@

\vspace{5mm}
\reply: Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c.

@@

\vspace{5mm}
\noindent \aecom \textbf{1 - Reply to AE comment 4}: 
I hank you for explaining the sampling limitations to me in your reply.  However, I think these limitations should also be noted in the narrative, and the sub-optimality of the resulting trajectory should be acknowledged.

\vspace{5mm}
\reply: We agree that this should be mentioned in the text. We have added the following text in the Closing remarks section at page 25.

\begin{quote}
    "The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources available on the AUV, assimilation into a discretized fixed grid is necessary.``
\end{quote}

@@

\vspace{5mm}
\reply: We have removed the word sampling from the sentence, with made it more clear and readable. See the change below:

\begin{quote}
    "Models and methods from spatial statistics and experimental design can clearly contribute to this challenge.``
\end{quote}

@@

\vspace{5mm}
\reply: Changed to “proxies” on page 4.

@@

\vspace{5mm}
\reply: We have changed the reference accordingly in the text at Page 5.

### 2021-01-05 main_new.tex
Updates from Overleaf

Monitoring the world's oceans has gained increased importance in light
of the changing climate and increasing anthropogenic impact. Central
to understanding the changes taking place in the upper water-column is
knowledge of the bio-geophysical interaction driven by an
agglomeration of physical forcings (e.g. wind, topography, bathymetry,
tidal influences, etc.) and incipient micro-biology driven by
planktonic and coastal anthropogenic input, such as pollution and
agricultural runoff transported into the ocean by rivers.  These often
result in a range of ecosystem-related phenomena such as blooms and
plumes, with direct and indirect effects on society \citep{ryan2017}. One of the
bottlenecks in the study of such phenomena lies however in the lack of
observational data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations in which ocean processes transpire, prompting the need for
effective means of data collection.  By \emph{sampling}, we refer here
primarily to the design of observational strategies in the spatial
domain with the aim to pursue measurements with high scientific
relevance.  Models and methods from spatial statistics and experimental design can clearly contribute to this sampling challenge.

@@

The work presented here is primarily inspired by a case study
pertaining to using an AUV for spatial characterization of a frontal
system generated by a river plume. Fig.~\ref{fig:nidelven} shows the
survey area in Trondheim, Norway, where cold freshwater enters the fjord from a river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coriolis force the
cold fresh water tends to flow east. Depending on the variations in
river discharge, tidal effects, coastal current and wind, this
boundary often gets distorted, and knowledge about its location is
highly uncertain, making deterministic planning challenging. The goal
is therefore to use AUV measurements for improved description of the
interface between fresh and oceanic waters.  It is often not possible
to sample the biological variables of fundamental interest in such AUV
operations, but off-the-shelf instruments provide temperature and
salinity measurements which serve as proxys for the underlying
biological phenomenon. With the help of a vector-valued random field
model for temperature and salinity, one can then aim to describe the
plume.  The goal of plume characterization, in this way, relates to
that of estimating some regions of the domain, typically excursion
sets (ESs), when implicitly defined by the vector-valued random field.
In our context of environmental sampling, the joint salinity and
temperature excursions of a river plume help characterize the
underlying bio-geochemical processes
\citep{hopkins2013detection,Pinto2018}. Motivating examples for ESs of
multivariate processes are also abundant in other contexts, for
instance in medicine, where physicians do not rely solely on a single
symptom but must see several combined effects before making a
diagnosis.

@@

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, the PhD thesis \citep[][p.82]{stroh} mentions general
possibilities of stepwise uncertainty reduction strategies for ES's in
the context of designing fire simulations, yet outputs are mainly
assumed independent.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.

>>>>

Monitoring the world's oceans has gained increased importance in light
of the changing climate and increasing anthropogenic impact. Central
to understanding the changes taking place in the upper water-column is
knowledge of the bio-geophysical interaction driven by an
agglomeration of physical forcings (e.g. wind, topography, bathymetry,
tidal influences, etc.) and incipient micro-biology driven by
planktonic and coastal anthropogenic input, such as pollution and
agricultural runoff transported into the ocean by rivers.  These often
result in a range of ecosystem-related phenomena such as blooms and
plumes, with direct and indirect effects on society \citep{ryan2017}. One of the
bottlenecks in the study of such phenomena lies however in the lack of
observational data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations in which ocean processes transpire, prompting the need for
effective means of data collection.  By \emph{sampling}, we refer here
primarily to the design of observational strategies in the spatial
domain with the aim to pursue measurements with high scientific
relevance. Models and methods from spatial statistics and experimental design can clearly contribute to this challenge.

@@

The work presented here is primarily inspired by a case study
pertaining to using an AUV for spatial characterization of a frontal
system generated by a river plume. Fig.~\ref{fig:nidelven} shows the
survey area in Trondheim, Norway, where cold freshwater enters the fjord from a river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coriolis force the
cold fresh water tends to flow east. Depending on the variations in
river discharge, tidal effects, coastal current and wind, this
boundary often gets distorted, and knowledge about its location is
highly uncertain, making deterministic planning challenging. The goal
is therefore to use AUV measurements for improved description of the
interface between fresh and oceanic waters.  It is often not possible
to sample the biological variables of fundamental interest in such AUV
operations, but off-the-shelf instruments provide temperature and
salinity measurements which serve as proxies for the underlying
biological phenomenon. With the help of a vector-valued random field
model for temperature and salinity, one can then aim to describe the
plume.  The goal of plume characterization, in this way, relates to
that of estimating some regions of the domain, typically excursion
sets (ESs), when implicitly defined by the vector-valued random field.
In our context of environmental sampling, the joint salinity and
temperature excursions of a river plume help characterize the
underlying bio-geochemical processes
\citep{hopkins2013detection,Pinto2018}. Motivating examples for ESs of
multivariate processes are also abundant in other contexts, for
instance in medicine, where physicians do not rely solely on a single
symptom but must see several combined effects before making a
diagnosis.

@@

Surprisingly less attention has been dedicated to sequential
strategies in the case of vector-valued observations. It has been long
acknowledged that co-Kriging could be updated efficiently in the context of sequential data assimilation \citep{Vargas-Guzman1999}, but sequential
strategies for estimating features of vector-valued random fields are
still in their infancy. \cite{LeGratiet.etal2015} used co-Kriging
based sequential designs to multi-fidelity computer codes and
\cite{Poloczek2017} used related ideas for multi-information source
optimization, but not for ES's like we do here. More relevant to our
setting, \citet[p.82]{stroh} mentions general
possibilities of step-wise uncertainty reduction strategies for ES's in
the context of designing fire simulations, yet outputs are mainly
assumed independent.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is necessary. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.
